Notes on Wide Gap

### 2016-05-16 paper/wide_gap_notes.tex
finer grid search for Pm=1E-2
N31 term added
weakly nonlinear Latter Fromang and Faure eqns

The purely conducting boundary condition states that the axial component of the current ($\mathbf{J}_z = [\nabla \times \mathbf{B}]_z)$ must be zero at the walls. In the thin gap geometry, the purely conducting boundary condition on the azimuthal magnetic field is $\partial_x(B_y) = 0$ for axisymmetric perturbations. A spatially constant azimuthal field satisfies both the thin-gap MRI equations and this boundary condition. This neutral mode is formally included in the analysis of Umurhan et al 2007 and yields a second amplitude equation in the form of a simple diffusion equation. This amplitude equation decouples from the GLE because of the translational symmetry of the thin-gap geometry. Because that symmetry is not preserved in the wide-gap case, Umurhan et al postulate that slow variation in the wide-gap geometry will be governed by two coupled amplitude equations. However, the wide-gap conducting boundary condition $J_z = \frac{1}{r} \partial_r (r B_\phi) = 0$, together with the purely geometric term in Equation \ref{eq:By_perturbed}, conspire to prevent the wide-gap geometry from sustaining a neutral mode.

-- coefficients of GLE come from inner products, therefore must always be real. Could we ever get a complex GLE?

@@

\begin{figure}[h!]
\centering
\includegraphics[width=\textwidth]{../../weakly_nonlinear_mri/python/widegap/figures/u_near_threshold_test_critical}
\caption{Wide-gap first-order velocity perturbations near threshold. Unnormalized. This is with Rm = 0.845, Q = 0.006.
\label{fig:firstorder_u}}
\end{figure}

\begin{figure}[h!]
\centering
\includegraphics[width=\textwidth]{../../weakly_nonlinear_mri/python/widegap/figures/B_near_threshold_test_critical}
\caption{Wide-gap first-order B-field perturbations near threshold. Unnormalized. This is with Rm = 0.845, Q = 0.006.
\label{fig:firstorder_B}}
\end{figure}

\begin{figure}[h!]
\centering
\includegraphics[width=\textwidth]{../../weakly_nonlinear_mri/python/widegap/figures/firstorderperturbations_test}
\caption{First order perturbations, Rm = 0.845, Q = 0.006.
\label{fig:firstorder_all}}
\end{figure}

\begin{figure}[h!]
\centering
\includegraphics[width=\textwidth]{../../weakly_nonlinear_mri/python/widegap/figures/N2_test}
\caption{N2 from unnormalized lower-order components. This is with Rm = 0.845, Q = 0.006. Black = real, teal = imaginary. Why is psi not zero at the boundaries? Assuming it's because of the additional B advective term, not present in the thin-gap.
\label{fig:N2}}
\end{figure}

>>>>

The purely conducting boundary condition states that the axial component of the current ($\mathbf{J}_z = [\nabla \times \mathbf{B}]_z)$ must be zero at the walls. In the thin gap geometry, the purely conducting boundary condition on the azimuthal magnetic field is $\partial_x(B_y) = 0$ for axisymmetric perturbations. A spatially constant azimuthal field satisfies both the thin-gap MRI equations and this boundary condition. This neutral mode is formally included in the analysis of Umurhan et al 2007 and yields a second amplitude equation in the form of a simple diffusion equation. This amplitude equation decouples from the GLE because of the translational symmetry of the thin-gap geometry. Because that symmetry is not preserved in the wide-gap case, Umurhan et al postulate that slow variation in the wide-gap geometry will be governed by two coupled amplitude equations. However, the wide-gap conducting boundary condition $J_z = \frac{1}{r} \partial_r (r B_\phi) = 0$, together with the purely geometric term in Equation \ref{eq:By_perturbed}, conspire to prevent the wide-gap geometry from sustaining a neutral mode. We note that this mode would exist in a resistance-free approximation.

The nonlinear terms in the wide-gap case differ from their thin-gap analogues mainly in the addition of vertical advective terms. These terms derive from the advective derivatives in the momentum and induction equations, but are filtered out in the thin-gap approximation. These terms have several interesting consequences. First, they allow the nonlinear contributions at both second and third order (i.e. N$_2$ and N$_3$) not to individually satisfy the boundary conditions on $\Psi$ and $u$. Second, the advection terms break the alternating parity exhibited by the thin-gap nonlinear terms.

@@

\begin{figure}[h!]
\centering
\includegraphics[width=\textwidth]{../../weakly_nonlinear_mri/python/widegap/figures/u_near_threshold_test_critical}
\caption{Wide-gap first-order velocity perturbations near threshold. Unnormalized. This is with Pm = 1E-3, Rm = 0.845, Q = 0.006.
\label{fig:firstorder_u}}
\end{figure}

\begin{figure}[h!]
\centering
\includegraphics[width=\textwidth]{../../weakly_nonlinear_mri/python/widegap/figures/B_near_threshold_test_critical}
\caption{Wide-gap first-order B-field perturbations near threshold. Unnormalized. This is with Pm = 1E-3, Rm = 0.845, Q = 0.006.
\label{fig:firstorder_B}}
\end{figure}

\begin{figure}[h!]
\centering
\includegraphics[width=\textwidth]{../../weakly_nonlinear_mri/python/widegap/figures/firstorderperturbations_test}
\caption{First order perturbations, Pm = 1E-3, Rm = 0.845, Q = 0.006.
\label{fig:firstorder_all}}
\end{figure}

\begin{figure}[h!]
\centering
\includegraphics[width=\textwidth]{../../weakly_nonlinear_mri/python/widegap/figures/N2_test}
\caption{N2 from unnormalized lower-order components. This is with Pm = 1E-3, Rm = 0.845, Q = 0.006. Black = real, teal = imaginary. Why is psi not zero at the boundaries? Assuming it's because of the additional B advective term, not present in the thin-gap.
\label{fig:N2}}
\end{figure}

\begin{figure}[h!]
\centering
\includegraphics[width=\textwidth]{../../weakly_nonlinear_mri/python/widegap/figures/n3_test}
\caption{N31 from unnormalized lower-order components. This is with Pm = 1E-3, Rm = 0.845, Q = 0.006. Black = real, teal = imaginary. 
\label{fig:N31}}
\end{figure}
